---
title: "Track-a-Cohort: Winter-run Chinook Figures"
date: "Last updated: `r Sys.Date()`"
format: html
output:
  html_document: 
   toc: true
   toc_float: true
   code_folding: true
   cache: true 
editor: visual
bibliography: ../www/refs/references.bib
---

## Background

Objective: Using shared resources via [BDO github](https://github.com/BDO-Science/track-a-cohort) from BOR, replicate figures and adjust to include dynamic data. See [Track a cohort_WR_plots.docx](https://github.com/BDO-Science/track-a-cohort/blob/main/Track%20a%20Cohort_WR_plots.docx) for figures requested. Also consider future additions including more interactive capabilities to certain graphs in form of Rmarkdown and/or shinyapp

```{r, include=FALSE}
library(here)
source(here("R/load_dependencies.R"))

# extract current year
current_year<-year(today())
```

## Figure 1. Juvenile Production Estimate

```{r, load_plot_01, fig.width=10, fig.height=6, echo=FALSE}
source(here("R/winter_run_chinook_plot_annual_jpe.R"))
```

## Figure 2. Survival and Routing Probabilites for Winter-run Chinook

```{r, load_plot_02, fig.width=12, fig.height=12, echo=FALSE}
source(here("R/winter_run_chinook_plot_STARS.R"))
```

## Figure 3. Juvenile Production Estimate (JPE) – Percent Loss

### Figure 3a. Genetic vs Length-At-Date (LAD) Historical Percent Loss of the JPE

```{r load_plot_03a, fig.width=10, fig.height=6, echo=FALSE}
source(here("R/winter_run_chinook_plot_pct_genetic_lad_loss_jpe.R"))
```

### Figure 3b. Cumulative percent loss of the JPE

```{r, load_plot_03b, fig.width=10, fig.height=6, echo=FALSE, warning = FALSE}
source(here("R/winter_run_chinook_plot_pct_cumulative_genetic_loss_jpe.R"))
```

## Figure 4. Juvenile Production Estimate (JPE) – Number Loss

### Figure 4a. Genetic vs Length-At-Date (LAD) Historical Number Loss 

### Figure 4b. 
