---
title: "Track-a-Cohort: Steelhead"
date: "Last updated: `r Sys.Date()`"
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

## Background

This document uses shared resources via [BDO github](https://github.com/BDO-Science/track-a-cohort){target="_blank"} from BOR to replicate figures requested and adjust underlying code to include dynamic data. See [Track a cohort_Steelhead.docx](https://github.com/BDO-Science/track-a-cohort/blob/main/SH_Cohort/Track%20a%20Cohort_Steelhead.docx){target="_blank"} for figures requested. Certain figures include a link to more interactive plot types using Shiny (in development) and all figures include a link to code in separate CBR developed [github repo](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort){target="_blank"}.

```{r, include=FALSE}
library(here)
source(here("R/load_dependencies.R"))

#set current year
source(here("R/utils_fct_assign_current_water_year.R"))
current_year <- assign_current_water_year()
```

## RBDD Juvenile Passage Estimates

```{r, load_plot_rbdd_biweekly_passage, fig.width=10, fig.height=7}
source(here("R/steelhead_plot_rbdd_biweekly_passage.R"))
```

Related links: [SacPAS Page](https://www.cbr.washington.edu/sacramento/tmp/redbluffdaily_1714603318_94.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/winter_run_chinook_plot_annual_jpe.R){target="_blank"}

-   Issues:
    -   Remove from R code? If keeping:
    -   Update biweekly shaded areas based on SI generated methods
    -   Continue to match the same 0 line cutoff as on SacPAS or leave space to show points at 0 line?
    -   Confirm Calendar Year or Water Year. Currently code only pulls current year based on today's date. To look at water year will need to adjust code to include query link.

## Delta STARs Survival and Routing Probabilities

Note: Plots below use Winter-run Chinook as a surrogate for Steelhead

```{r,  load_plot_STARS, fig.width=12, fig.height=8}
source(here("R/winter_run_chinook_plot_STARS.R"))
```

Related links: [Interactive Plot](https://caitobrien.shinyapps.io/track-a-cohort_WRChinookSTARS/){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/winter_run_chinook_plot_STARS.R){target="_blank"}, [STARS ShinyApp](https://www.cbr.washington.edu/shiny/STARS/){target="_blank"}

## Total Loss - Clipped and Unclipped Steelhead

### 

```{r load_plot_count_total_loss, fig.width=10, fig.height=7}
source(here("R/steelhead_plot_count_total_loss.R"))
```

Related links: [SacPAS Query](https://www.cbr.washington.edu:2024/sacramento/data/query_loss_detail.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_count_total_loss.R){target="_blank"}

Issues:

-   Confirm data source footnote

## Size (Fork Length) Distribution of Steelhead Loss

```{r load_plot_size_distribution_loss, fig.width=10, fig.height=7, fig.cap="Figure compares density plots that highlight historical (WY1994 to WY2023) size distribution and histogram of current year (WY2024) size distribution by rear type (yellow = unclipped; blue = clipped). Fork lengths below 750 mm were included in dataset."}
source(here("R/steelhead_plot_size_distribution_loss.R"))
```

Related links: [SacPAS Query](https://www.cbr.washington.edu:2024/sacramento/data/query_loss_detail.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_size_distribution_loss.R){target="_blank"}

```{r load_plot_size_distribution_by_year_loss, fig.width=8, fig.height=10}
source(here("R/steelhead_plot_size_distribution_by_year_loss.R"))
```

Related links: [SacPAS Query](https://www.cbr.washington.edu:2024/sacramento/data/query_loss_detail.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_size_distribution_by_year_loss.R){target="_blank"}

## Cumulative Loss with Single-Year Thresholds

```{r load_plot_count_cumulative_loss_single_year_thresholds, fig.width=10, fig.height=8}
source(here("R/steehead_plot_count_cumulative_loss_single_year_thresholds.R"))
```

Related links: [SacPAS Page](https://www.cbr.washington.edu/sacramento/workgroups/salmon_monitoring.html#loss:~:text=Chinook%20Salmon%20Loss-,Natural%20Central%20Valley%20Steelhead,-Daily%20Loss%20Table){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steehead_plot_count_cumulative_loss_single_year_thresholds.R){target="_blank"}

Issues:

-   Confirm Single-Year-Thresholds and update to generate automatically
    -   Tillotson etl al. 2022: "Because less information is available on annual wild Central Valley Steelhead natal origin and abundance, the maximum ITL in a single year is fixed at a loss of 2,760 between December 1 and March 31, and a loss of 3,040 between April 1 and June 15 (NMFS 2019). "

    -   SacPAS: Single-Year Loss Thresholds (PA 4-69): In each year, typically January/February, Reclamation and DWR propose to avoid exceeding an annual loss threshold equal to 90% of the greatest annual loss that occurred in the historical record 2010-2018 for each of:

        -   Natural Winter-Run Chinook Salmon (loss= 1.17% of JPE)

        -   Natural Central Valley Steelhead from December through March (loss =1,414)

        -   Natural Central Valley Steelhead from April through June 15 (loss = 1,552)

        (More information on 4-70)
-   Confirm only including unclipped
- ShinyApp pending

## Cumulative Loss by BiOp Status and Hydrological Classification Index (HCI)

```{r load_plot_count_cumulative_loss, fig.width=10, fig.height=10, fig.cap="The figure shows cumulative loss by BiOp Status and Hydrological Classification Index (HCI). Each quadrant of the faceted plot includes grey lines for historical years, colored lines (blue for wet years, red for dry years) for years within the BiOp status and HCI type, a black line for the current year, and a dashed horizontal line indicating the current cumulative loss maximum."}
source(here("R/steelhead_plot_count_cumulative_loss.R"))
```

Related links: [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_count_cumulative_loss.R){target="_blank"}

Issues:

-   Confirm use of clipped and unclipped in dataset

-   Add legend key

-   Confirm data source footnote

-   ShinyApp pending

## Daily Loss and Export by Pumping Facility

```{r, load_plot_daily_loss_export, fig.width=12, fig.height=10}
source(here("R/steelhead_plot_daily_loss_export.R"))
```

Related links: SacPAS, [Interactive Plot - ShinyApp](https://caitbrien.shinyapps.io/track-a-cohort_DailyLossExports/){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_daily_loss_export.R){target="_blank"}

-   Issues:
    -   Confirm if OMRI vertical bars are designated by value or dates, is there a rule to be applied? Shared code is static values, but notes within [assessment pdf](chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://www.usbr.gov/mp/bdo/docs/20240528-proposed-action-assessment.pdf) states: "Vertical black bars are approximate dates when OMRI controlling factors change."

    -   Update ShinyApp

## Predicted Weekly Loss - Tillotson Model

```{r, load_plot_tillotson_predicted_loss, fig.width=10, fig.height=7}
source(here("R/steelhead_plot_tillotson_predicted_loss.R"))
```

Related links: SacPAS Page, [SacPAS Tillotson Tool](https://www.cbr.washington.edu/sacramento/lossandsalvage/){target="_blank"}, GitHub Repo Code: [Tillotson model](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/Steelhead_Model_Setup.R){target="_blank"}, [data wrangling and prediction output](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/utils_fct_predict_tillotson_model.R){target="_black"}, [plot output](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_tillotson_predicted_loss.R){target="_blank"}

-   Issues:
    -   Currently using BOR supplied code to run model, confirm same output with NB code.
        -   Update: NB believes this is updated code and will look into comparing and update Loss and Salvage Predictor Tool as needed when time allows (Estimates time in August).
    -   Confirm shared code is duplicate of the most up-to-date Tillotson code. If this is Tillotson code confirm with authors on use and confirm permissions to include model code in public facing repo. Alternatively, pull results from Loss and Salvage Predictor Tool.
        -   JG or NB to reach out?
    -   Confirm change in plot design

```{r, load_table_predicted_observed_loss, fig.width=10, fig.height=6}
# Load the data
load(here("data/tillotson_prediction_output.rda"))

data <- tillotson_prediction_output$steelhead_tillotson_output

#load function to wrangle into table format
source(here("R/utils_fct_generate_table_predicted_observed_loss.R"))

# generate HTML table
generate_table_pred_obs_loss(data, species = "Steelhead")

```

Related links: SacPAS Page,[SacPAS Tillotson Tool](https://www.cbr.washington.edu/sacramento/lossandsalvage/){target="_blank"}, GitHub Repo Code: [Tillotson model](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/Steelhead_Model_Setup.R){target="_blank"}, [data wrangling and prediction output](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/utils_fct_predict_tillotson_model.R){target="_black"}, [Table configuration](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/utils_fct_generate_table_predicted_observed_loss.R){target="_blank"}

-   Issues:
    -   See tillotson figure for issues
