---
title: "Track-a-Cohort: Steelhead"
date: "Last updated: `r Sys.Date()`"
format: html
output:
  html_document: 
   toc: true
   toc_float: true
   code_folding: true
   cache: true 
editor: visual
bibliography: ../www/refs/references.bib
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

## Background

This document uses shared resources via [BDO github](https://github.com/BDO-Science/track-a-cohort){target="_blank"} from BOR to replicate figures requested and adjust underlying code to include dynamic data. See [Track a cohort_Steelhead.docx](https://github.com/BDO-Science/track-a-cohort/blob/main/SH_Cohort/Track%20a%20Cohort_Steelhead.docx){target="_blank"} for figures requested. Certain figures include a link to more interactive plot types using Shiny (in development) and all figures include a link to code in separate CBR developed [github repo](https://github.com/caitobrien/TAC){target="_blank"}.

```{r, include=FALSE}
library(here)
source(here("R/load_dependencies.R"))

#set current year
source(here("R/utils_fct_assign_current_water_year.R"))
current_year <- assign_current_water_year()
```

## Figure 1. RBDD Juvenile Passage Estimates

```{r, load_plot_01, fig.width=10, fig.height=6}
source(here("R/steelhead_plot_rbdd_biweekly_passage.R"))
```

Related links: [SacPAS Page](https://www.cbr.washington.edu/sacramento/tmp/redbluffdaily_1714603318_94.html){target="_blank"}, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/winter_run_chinook_plot_annual_jpe.R){target="_blank"}

-   Issues:
    -   Update biweekly shaded areas based on SI generated methods
    -   Continue to match the same 0 line cutoff as on SacPAS or leave space to show points at 0 line?
    -   Confirm Calendar Year or Water Year. Currently code only pulls current year based on today's date. To look at water year will need to adjust code to include query link.

## Figure 2. Survival and Routing Probabilites

Note: Plots below use Winter-run Chinook as a surrogate for Steelhead

### Option 1: Facetted by year

```{r, load_plot_02, fig.width=12, fig.height=18 }
source(here("R/winter_run_chinook_plot_STARS.R"))
```

### Option 2: Without faceted years

```{r,  load_plot_02_alt, fig.width=12, fig.height=18}
source(here("R/winter_run_chinook_plot_STARS_withoutfacetedyears.R"))
```

Related links: SacPAS Page, [Interactive Plot](https://csobrien.shinyapps.io/WRChinookSTARS/){target="_blank"}, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/winter_run_chinook_plot_STARS.R){target="_blank"}, [STARS ShinyApp](https://www.cbr.washington.edu/shiny/STARS/){target="_blank"}

## Figure 3. Comparison of Clipped and Unclipped Loss

### Figure 3a. Total Loss of Clipped and Unclipped Steelhead

```{r load_plot_03a, fig.width=10, fig.height=6}
source(here("R/steelhead_plot_count_total_loss.R"))
```

Related links: SacPAS Page, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/steelhead_plot_count_total_loss.R){target="_blank"}

### Figure 3b. Size Distribution of Loss

```{r load_plot_03b, fig.width=10, fig.height=6, fig.cap="Figure compares density plots that highlight historical (WY1994 to WY2023) size distribution and histogram of current year (WY2024) size distribution by rear type (dark grey = unclipped; light grey = clipped). Fork lengths below 750 mm were included in dataset."}
source(here("R/steelhead_plot_size_distribution_loss.R"))
```

Related links: SacPAS Page, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/steelhead_plot_size_distribution_loss.R){target="_blank"}

### Figure 3c. Size Distribution of Steelhead Loss By Year

```{r load_plot_03c, fig.width=6, fig.height=10}
source(here("R/steelhead_plot_size_distribution_by_year_loss.R"))
```

Related links: SacPAS Page, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/steelhead_plot_size_distribution_by_year_loss.R){target="_blank"}

## Figure 4. Cumulative Loss

```{r load_plot_04, fig.width=10, fig.height=10, fig.cap="The figure shows cumulative loss by BiOp Status and Hydrological Classification Index (HCI). Each quadrant of the faceted plot includes grey lines for historical years, colored lines (blue for wet years, red for dry years) for years within the BiOp status and HCI type, a black line for the current year, and a dashed horizontal line indicating the current cumulative loss maximum."}
source(here("R/steelhead_plot_count_cumulative_loss.R"))
```

Related links: SacPAS Page, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/steelhead_plot_count_cumulative_loss.R){target="_blank"}, [Interactive Plot - ShinyApp](https://csobrien.shinyapps.io/CumulativeLoss/){target="\"_blank"}

## Figure 5. Cumulative Loss with Single-Year Thresholds

```{r load_plot_05, fig.width=10, fig.height=6}
source(here("R/steehead_plot_count_cumulative_loss_single_year_thresholds.R"))
```

Related links: [SacPAS Page](https://www.cbr.washington.edu/sacramento/workgroups/salmon_monitoring.html#loss:~:text=Chinook%20Salmon%20Loss-,Natural%20Central%20Valley%20Steelhead,-Daily%20Loss%20Table){target="_blank"}, ShinyApp, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/steehead_plot_count_cumulative_loss_single_year_thresholds.R){target="_blank"}, [Interactive Plot - ShinyApp](https://csobrien.shinyapps.io/CumulativeLossSingleYearThreshold/){target="\"_blank"}

-   Issues:
    -   Confirm Single-Year-Thresholds and update to generate automatically
    -   Confirm only including unclipped

## Figure 6. Daily Loss and Export by Pumping Facility

```{r, load_plot_06, fig.width=12, fig.height=10}
source(here("R/steelhead_plot_daily_loss_export.R"))
```

Related links: SacPAS Page, [Interactive Plot - ShinyApp](https://csobrien.shinyapps.io/DailyLossExports/){target="_blank"}, [GitHub Repo Code](https://github.com/caitobrien/TAC/blob/main/R/steelhead_plot_daily_loss_export.R){target="_blank"}

-   Issues:
    -   Confirm OMRI is value wanted - different from tillotson model. Based on shared code and description in Assessment pdf's OMRI is used for management.
    -   Confirm if OMR/OMRI vertical bars are designated by value or dates, is there a rule to be applied? Shared code is static values, but notes within [assessment pdf](chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://www.usbr.gov/mp/bdo/docs/20240528-proposed-action-assessment.pdf) states: "Vertical black bars are approximate dates when OMRI controlling factors change."

## Figure 7. Predicted and Observed Weekly Loss - Tillotson Model

```{r, load_plot6, fig.width=10, fig.height=6}
source(here("R/steelhead_plot_tillotson_predicted_loss.R"))
```

Related links: SacPAS Page, [SacPAS Tillotson Tool](https://www.cbr.washington.edu/sacramento/lossandsalvage/){target="_blank"}, GitHub Repo Code: [Tillotson model](https://github.com/caitobrien/TAC/blob/main/data-raw/Steelhead_Model_Setup.R){target="_blank"}, [data wrangling and prediction output](https://github.com/caitobrien/TAC/blob/main/data-raw/utils_fct_predict_tillotson_model.R){target="_black"}, [plot output](https://github.com/caitobrien/TAC/blob/main/R/steelhead_plot_tillotson_predicted_loss.R){target="_blank"}

- Issues:

  -   Currently using BOR supplied code to run model, confirm same output with NB code
  -   Further, confirm shared code is duplicate of Tillotson code. If not, update to reflect Tillotson model
  -   Confirm mean() is used for all averages within tillotson code and not rolling average.
  -   Confirm change in plot design

## Table 1. Model inputs and predictions - Tillotson Model

```{r, load_table_01, fig.width=10, fig.height=6}
# Load the data
load(here("data/tillotson_prediction_output.rda"))

data <- tillotson_prediction_output$steelhead_tillotson_output

#load function to wrangle into table format
source(here("R/utils_fct_generate_table_predicted_observed_loss.R"))

# generate HTML table
generate_table_pred_obs_loss(data, species = "Steelhead")

```

Related links: SacPAS Page,[SacPAS Tillotson Tool](https://www.cbr.washington.edu/sacramento/lossandsalvage/){target="_blank"}, GitHub Repo Code: [Tillotson model](https://github.com/caitobrien/TAC/blob/main/data-raw/Steelhead_Model_Setup.R){target="_blank"}, [data wrangling and prediction output](https://github.com/caitobrien/TAC/blob/main/data-raw/utils_fct_predict_tillotson_model.R){target="_black"} [Table configuration](https://github.com/caitobrien/TAC/blob/main/R/utils_fct_generate_table_predicted_observed_loss.R){target="_blank"}

-   Issues:
    -   Confirm model output weekly averages. Table/Model is using weekly averages with mean(), confirm paper doesn't use a rolling average.
