---
title: "Track-a-Cohort: Steelhead"
date: "Last updated: `r Sys.Date()`"
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

## Background

This document uses shared resources via [BDO github](https://github.com/BDO-Science/track-a-cohort){target="_blank"} from BOR to replicate figures requested and adjust underlying code to include dynamic data. See [Track a cohort_Steelhead.docx](https://github.com/BDO-Science/track-a-cohort/blob/main/SH_Cohort/Track%20a%20Cohort_Steelhead.docx){target="_blank"} for figures requested. Certain figures include a link to more interactive plot types using Shiny (in development) and all figures include a link to code in separate CBR developed [github repo](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort){target="_blank"}.

```{r, include=FALSE}
library(here)
source(here("R/load_dependencies.R"))

#set current year
source(here("R/utils_fct_assign_current_water_year.R"))
current_year <- assign_current_water_year()
```

## RBDD Juvenile Passage Estimates

```{r, load_plot_rbdd_biweekly_passage, fig.width=10, fig.height=7}
source(here("R/steelhead_plot_rbdd_biweekly_passage.R"))
```

Related links: [SacPAS Page](https://www.cbr.washington.edu/sacramento/tmp/redbluffdaily_1714603318_94.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/winter_run_chinook_plot_annual_jpe.R){target="_blank"}

## Delta STARs Survival and Routing Probabilities

Note: Plots below use Winter-run Chinook as a surrogate for Steelhead

```{r,  load_plot_STARS, fig.width=12, fig.height=8}
source(here("R/winter_run_chinook_plot_STARS.R"))
```

Related links: [Interactive Plot](https://caitobrien.shinyapps.io/track-a-cohort_WRChinookSTARS/){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/winter_run_chinook_plot_STARS.R){target="_blank"}, [STARS ShinyApp](https://www.cbr.washington.edu/shiny/STARS/){target="_blank"}

## Total Loss - Clipped and Unclipped Steelhead

### 

```{r load_plot_count_total_loss, fig.width=10, fig.height=7}
source(here("R/steelhead_plot_count_total_loss.R"))
```

Related links: [SacPAS Query](https://www.cbr.washington.edu:2024/sacramento/data/query_loss_detail.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_count_total_loss.R){target="_blank"}

## Size (Fork Length) Distribution of Steelhead Loss

```{r load_plot_size_distribution_loss, fig.width=10, fig.height=7, fig.cap="Figure compares density plots that highlight historical (WY1994 to WY2023) size distribution and histogram of current year (WY2024) size distribution by rear type (yellow = unclipped; blue = clipped). Fork lengths below 750 mm were included in dataset."}
source(here("R/steelhead_plot_size_distribution_loss.R"))
```

Related links: [SacPAS Query](https://www.cbr.washington.edu:2024/sacramento/data/query_loss_detail.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_size_distribution_loss.R){target="_blank"}

```{r load_plot_size_distribution_by_year_loss, fig.width=8, fig.height=10}
source(here("R/steelhead_plot_size_distribution_by_year_loss.R"))
```

Related links: [SacPAS Query](https://www.cbr.washington.edu:2024/sacramento/data/query_loss_detail.html){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_size_distribution_by_year_loss.R){target="_blank"}

## Cumulative Loss with Single-Year Thresholds

```{r load_plot_count_cumulative_loss_single_year_thresholds, fig.width=10, fig.height=8}
source(here("R/steehead_plot_count_cumulative_loss_single_year_thresholds.R"))
```

Related links: [SacPAS Page](https://www.cbr.washington.edu/sacramento/workgroups/salmon_monitoring.html#loss:~:text=Chinook%20Salmon%20Loss-,Natural%20Central%20Valley%20Steelhead,-Daily%20Loss%20Table){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steehead_plot_count_cumulative_loss_single_year_thresholds.R){target="_blank"}

## Cumulative Loss by BiOp Status and Hydrological Classification Index (HCI)

```{r load_plot_count_cumulative_loss, fig.width=10, fig.height=10, fig.cap="The figure shows cumulative loss by BiOp Status and Hydrological Classification Index (HCI). Each quadrant of the faceted plot includes grey lines for historical years, colored lines (blue for wet years, red for dry years) for years within the BiOp status and HCI type, a black line for the current year, and a dashed horizontal line indicating the current cumulative loss maximum."}
source(here("R/steelhead_plot_count_cumulative_loss.R"))
```

Related links: [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_count_cumulative_loss.R){target="_blank"}

## Daily Loss and Export by Pumping Facility

```{r, load_plot_daily_loss_export, fig.width=12, fig.height=10}
source(here("R/steelhead_plot_daily_loss_export.R"))
```

Related links: SacPAS, [Interactive Plot - ShinyApp](https://caitobrien.shinyapps.io/track-a-cohort_DailyLossExports/){target="_blank"}, [GitHub Repo Code](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_daily_loss_export.R){target="_blank"}

## Predicted Weekly Loss - Tillotson Model

```{r, load_plot_tillotson_predicted_loss, fig.width=10, fig.height=7}
source(here("R/steelhead_plot_tillotson_predicted_loss.R"))
```

Related links: SacPAS Page, [SacPAS Tillotson Tool](https://www.cbr.washington.edu/sacramento/lossandsalvage/){target="_blank"}, GitHub Repo Code: [Tillotson model](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/Steelhead_Model_Setup.R){target="_blank"}, [data wrangling and prediction output](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/utils_fct_predict_tillotson_model.R){target="_black"}, [plot output](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/steelhead_plot_tillotson_predicted_loss.R){target="_blank"}

```{r, load_table_predicted_observed_loss, fig.width=10, fig.height=6}
# Load the data
load(here("data/tillotson_prediction_output.rda"))

data <- tillotson_prediction_output$steelhead_tillotson_output

#load function to wrangle into table format
source(here("R/utils_fct_generate_table_predicted_observed_loss.R"))

# generate HTML table
generate_table_pred_obs_loss(data, species = "Steelhead")

```

Related links: SacPAS Page,[SacPAS Tillotson Tool](https://www.cbr.washington.edu/sacramento/lossandsalvage/){target="_blank"}, GitHub Repo Code: [Tillotson model](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/Steelhead_Model_Setup.R){target="_blank"}, [data wrangling and prediction output](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/data-raw/utils_fct_predict_tillotson_model.R){target="_black"}, [Table configuration](https://github.com/Columbia-Basin-Research-CBR/track-a-cohort/blob/main/R/utils_fct_generate_table_predicted_observed_loss.R){target="_blank"}
