---
title: "Track-a-Cohort: Steelhead Figures"
date: "Last updated: `r Sys.Date()`"
format: html
output:
  html_document: 
   toc: true
   toc_float: true
   code_folding: true
   cache: true 
editor: visual
bibliography: ../www/refs/references.bib
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

## Background

Objective: Using shared resources via [BDO github](https://github.com/BDO-Science/track-a-cohort) from BOR, replicate figures and adjust to include dynamic data. See [Track a cohort_Steelhead.docx](https://github.com/BDO-Science/track-a-cohort/blob/main/SH_Cohort/Track%20a%20Cohort_Steelhead.docx) for figures requested. Also consider future additions including more interactive capabilities to certain graphs in form of Rmarkdown and/or shinyapp

```{r, include=FALSE}
library(here)
source(here("R/load_dependencies.R"))

#set current year
source(here("R/utils_fct_assign_current_water_year.R"))
current_year <- assign_current_water_year()
```

## Figure 1. RBDD Juvenile Passage Estimates Steelhead

```{r, load_plot_01, fig.width=10, fig.height=6}
source(here("R/steelhead_plot_rbdd_biweekly_passage.R"))
```

## Figure 2. Survival and Routing Probabilites

Note: Using Winter-run Chinook as a surrogate for Steelhead

### Facetted by year

```{r, load_plot_02, fig.width=12, fig.height=12 }
source(here("R/winter_run_chinook_plot_STARS.R"))
```

### Without faceted years

```{r,  load_plot_02_alt, fig.width=12, fig.height=12}
source(here("R/winter_run_chinook_plot_STARS_withoutfacetedyears.R"))
```

## Figure 3. Comparison of Clipped and Unclipped Steelhead Loss

### Figure 3a. Current and Historical Total Loss of Clipped and Unclipped Steelhead

```{r load_plot_03a, fig.width=10, fig.height=6}
source(here("R/steelhead_plot_count_total_loss.R"))
```

### Figure 3b. Size Distribution of Steelhead Loss

```{r load_plot_03b, fig.width=10, fig.height=6, fig.cap="Underlying density plot shows historical (WY1994 to WY2023) size distribution by rear type. Current year (WY2024) size distribution is shown as a histogram by rear type. Fork lengths below 750 mm were included"}
source(here("R/steelhead_plot_size_distribution_loss.R"))
```

### Figure 3c. Size Distribution of Steelhead Loss By Year

```{r load_plot_03c, fig.width=6, fig.height=10}
source(here("R/steelhead_plot_size_distribution_by_year_loss.R"))
```

## Figure 4. Cumulative Steelhead Loss

```{r load_plot_04, fig.width=10, fig.height=10}
source(here("R/steelhead_plot_count_cumulative_loss.R"))
```

## Table 1. Minimization of Cumulative Loss of Hatchery Steelhead

\*Pending steelhead hatchery release information

```{r load_table_01, results='asis'}
source(here("R/steelhead_table_summary_of_hatchery_loss.R"))

```
